### 2019-02-26 summary/paper-summary.tex
added the first lines of the paper summary and a sample set of the results

\documentclass{article}

\usepackage{hyperref}
\usepackage{xcolor}

\begin{document}

	\title{Leveraged Volume Sampling - Summary}
	\author{Ohad Ohayon}
	\pagenumbering{gobble}
	\maketitle
	\newpage

	\pagenumbering{arabic}

    \tableofcontents
    \newpage

    \section{Overview}
        This document summarises the contribution of the {\color{blue}\href{https://arxiv.org/abs/1802.06749}{\underline{paper}}} \textit{"Leveraged volume sampling for linear regression."}.
        It follows the structure of the paper itself, summarizing each of it's main sections.
        Finally, I will demonstrate my attempt at reproducing the results in the paper.

    \section{Introduction}
        \subsection{The Problem}
            Assume a linear regression problem where the feature vectors are the rows of an $n$ by $d$ full rank matrix $X=(x_{i}^T)$,
            the labels are the column vector $y$ and the optimal solution is $w^\ast=X^+y$.
            Define the average loss function of a solution $w$ to be $L(w)=frac{1}_{n}||Xw-y||_{2}^2$.
            In many real-world scenarios acquiring labels for some linear regression problems can be an expensive operation,
            although the regression features are readily available (as they may be just the parameters for an experiment).
            You may be accessible to the labels but you must limit the amount you request in order to deliver results within
            reasonable resource constraints.
            Assume the rows of such a sub-sampled problem to be $S$, the sub-sampled matrix and labels vector to be $X_{S}$
            and $y_{S}$ respectively and the sub-sampled solution to be $w_{S}=X_{S}^+y_{S}$.
            We'd like an effective algorithm to choose $S$, based on the features alone, whose corresponding solution $w_{S}$
            will hold the following properties with regard to the optimal solution $w^\ast$:
            \begin{itemize}
                \item $w_{S}$ is unbiased, i.e.: $E[w_{S}]=w^\ast$.
                \item The error for $w_{S}$ can be bounded by $L(w_{S}) < (1+\epsilon)L(w^\ast)$ for an $\epsilon$ of choice.
                \item Can be executed efficiently.
            \end{itemize}

\end{document}